Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-11-11 files/Chapter3/dynamicconvergence.tex
symbols
Overview, Chapter 3

In \cite{BV13}, a projection $P$ is defined so that $Pu \in V$. The error is
then split using this projection. Let $e(t) = Pu(t)- u_h(t)$ and $e_p(t) = u(t)
	- Pu(t)$ so that
\begin{eqnarray}
	e^h(t) = e_p(t) +e(t). \label{Error}
\end{eqnarray}

According to the authors of~\cite{BV13}, interpolation theory can be used to determine the error $e_p$. However to obtain an error estimate for $e$ is not as trivial. This is where the use of the Finite Element Analysis comes in.

To simplify the proof, the problem is divided into a semi-discrete problem and a fully discrete problem.

@@

In \eqref{error_e}, the authors divide the error $e$ into a semi-discrete term
and a fully discrete term. The two terms are investigated as seperate problems
in \cite{BV13}.\\

@@

\newtheorem*{DC_Thm5}{Theorem 2}
\begin{DC_Thm5}
	{Main Result}\\
	If $f\in C^{2}([0,T]4,X)$, then
	\begin{eqnarray*}
		\Vert u_{h}(t_{k})-u_{k}^{h}\Vert_{W}\ & \leq & 7T^{2}\tau^{2}\max\Vert u_{h}^{(4)}\Vert_{W}+7T\tau^{2}\max\Vert u_{h}'''\Vert_{W}\\
		& & + \sqrt{2K_{a}}\tau^{4}\max\Vert u_{h}'''\Vert_{W}
	\end{eqnarray*}
	for each $t\in(0,\ T)$.
\end{DC_Thm5}

@@

\subsection{Obtaining the error estimate}
In this section we will show how the error estimate in Theorem 2 is obtained by
the authors of \cite{BV13}.\\

The following two assumptions are made in \cite{BV13}.
\subsubsection*{Assumptions}
\begin{itemize}
	\item[] \textbf{B1} - The solution $u \in C(J,V)$ of Problem GVar$^h$ has the property that $(Pu) \in C^2(J)$.

	\item[] \textbf{B2} - There exists a subspace $H$ of $V$ and a positive integer $\alpha$ such that if $w \in H$, then $$\inf_{v\in S^{h}}||w-v||_V \leq \hat{C} h^\alpha|||w|||_H,$$ where $|||w|||_H$ is a norm or a semi-norm for $H$.
\end{itemize}

The projection $P$ in section \eqref{e_est} is defined in \cite{BV13} as
\begin{eqnarray*}
	b(u-Pu,v) = 0 \ \ \ \textrm{ for all } v \in V.
\end{eqnarray*}

From assumption \textbf{B1}, $Pu \in C^2(J)$ and by the definition $P$, $P$ is
a projection with respect to the inner product $b$.\\

Subtracting \eqref{GV_PB1} from \eqref{DC_E2} gives
\begin{eqnarray*}
	c(u''(t)-u''_h(t),v) + a(u'(t)-u'_h(t),v) + b(u(t)-u_h(t),v) = 0
\end{eqnarray*}

Using the definition of the error $e^h$ while also adding and subtracting
$Pu(t)$ in the bilinear form $b$, we can write
\begin{eqnarray*}
	c(e_h''(t),v) + a(e_h'(t),v) + b(u(t) - Pu(t),v) + b(Pu(t) +u_h(t),v)- = 0.
\end{eqnarray*}

It then follows from the definitions of the error $e$ and the projection $P$
that
\begin{eqnarray*}
	c(e''(t),v) + a(e'(t),v) + b(e(t),v) = 0
\end{eqnarray*}

This equation \eqref{DC_E3} is presented in \cite{BV13} as a proposition.
\newtheorem{DC_Prop1}{Proposition}
\begin{DC_Prop1}
	If $u$ is the solution of Problem GVar and $u$ satisfies assumption \textbf{B1}, then
	\begin{eqnarray}
		c(e_h''(t),v) + a(e_h'(t),v) + b(e(t),v) = 0 \ \ \ \textrm{ for all } v \in V.
	\end{eqnarray}
\end{DC_Prop1}

Using this proposition, the authors of \cite{BV13} derive a generalized error
estimate for the error $e = Pu - u_h$.

\newtheorem{lem_1}{Lemma}
\begin{lem_1}
	If $u$ is the solution of Problem GVar and $u$ satisfies assumption \textbf{B1}, then
	\begin{eqnarray}
		\Vert e(t)\Vert_{W} &\leq& \sqrt{2} \biggl( \Vert e(0) \Vert_W + 3T\Vert e'_h(0) \Vert_W + 3 \int_o^T \Vert e'_p(t)\Vert_W\nonumber \\
		&& + 3K_\alpha T \Vert e_h(0) \Vert_W + 3K_\alpha \int_0^T \Vert e_p(t) \Vert_W \biggr). \label{fundamental_estimate}
	\end{eqnarray}
\end{lem_1}

In chapter 2, it was explained that the models of this dissertation have weak
damping, i.e. $a = 0$. Therefore, the error estimate in Lemma 1 can be
simplified. This is given in \cite{BV13} as a corollary.

\newtheorem{col_1}{Corollary}
\begin{col_1}
	If $u$ is the solution of Problem GVar with $a = 0$ and $u$ satisfies assumption \textbf{B1}, then
	\begin{eqnarray}
		\Vert e_(t)\Vert_{W} \leq \sqrt{2} \Vert e(0) \Vert_W + 2T||e'_h(0)||_W + 4\sqrt{T}\max_{t \in [0,T]} \Vert e'_p(t) \Vert_W.
	\end{eqnarray}
\end{col_1}

\begin{proof}
	Assume that there is no damping other than the weak damping, i.e. $a = 0$. The the equation in proposition 1 becomes
	\begin{eqnarray*}
		c(e_h''(t),v) + b(e(t),v) = 0 \ \ \ \textrm{ for all } v \in V.
	\end{eqnarray*}

	If $v \in S^h[0,T]$, then 
	\begin{eqnarray*}
		c(e_h''(t),v(t)) + b(e(t),v(t)) = 0.
	\end{eqnarray*}

	Let $v'(t) = c(t)$, then 
	\begin{flalign}
		\frac{d}{dt}\left[\frac{1}{2}c(e,e)-\frac{1}{2}b(v,v)-c(e_h,v) \right] &= c(e',e) - b(e,v)- c(e''_h, v)-c(e_h',v')\nonumber\\
		&= c(e'_p,e). \label{DC_E4}
	\end{flalign}

	Choose $v(\tau) = 0$ for some $\tau \in [0,T]$, and integrate \eqref{DC_E4} between $0$ and $\tau$ to obtain
	\begin{eqnarray*}
		a
	\end{eqnarray*}

\end{proof}

\subsubsection{Semi-discrete problem}
The estimate \eqref{fundamental_estimate} in Lemma 1 is equivalent to
\begin{eqnarray*}
	\Vert u(t) - u_h(t) \Vert_{W} &\leq& \sqrt{2} \biggl( \Vert u(0) - u_h(0) \Vert_W + 3T\Vert e'_h(0) \Vert_W + 3 \int_o^T \Vert e'_p(t)\Vert_W\nonumber \\
	&& + 3K_\alpha T \Vert e_h(0) \Vert_W + 3K_\alpha \int_0^T \Vert e_p(t) \Vert_W \biggr).
\end{eqnarray*}

From Theorem 1, it follows that the solution $u$ of Problem GVar satisfies $u
	\in C^1([0,T]; V) \cap C^2((0,T);V)$. \textcolor{red}{This is a weak assumption
	made in \cite{BV13}.}\\

\sout{The following is a weak assumption in \cite{BV13}: The solution $u$ of Problem GVar satisfies $u \in C^1([0,T]; V) \cap C^2((0,T);V)$. This follows from Theorem 1 and is a necessary assumption in [BSV17].}\\

Let $\Pi$ denote the interpolation operator such that $u_0^h = \Pi u_0$ and
$u_1^h = \Pi u_1$.

\newtheorem*{DC_Thm4}{Theorem 2}
\begin{DC_Thm4}
	Suppose Assumption E2 holds and $u_{0}^{h}=\Pi u_{0}$ and $u_{1}^{h}=\Pi u_{1}.$ If the solution $u$ of Problem $GVar$ satisfies Assumption El, $u(t)\in H$ and $u'(t)\in H$, then
	\begin{eqnarray*}
		\Vert u(t)-u_{h}(t)\Vert_{W} & \leq & C_{b} \hat{C}h^{\alpha}|||u(t)|||_{H}+\sqrt{2}C_{b}\hat{C}h^{\alpha}(3T\max|||u'(t)|||_{H}\\
		& & +3K_{a}T\max|||u(t)|||_{H}+(2+3K_{a}T)|||u_{0}|||_{H} \\
		& & +3T|||u_{1}|||_{H}) ,
	\end{eqnarray*} for each $t\in[0, T].$
\end{DC_Thm4}

\subsection{Fully discrete Galerkin finite element approximation}
Consider the time interval $J = [0,T]$. Divide $J$ into $N$ steps with length
$\tau = \frac{T}{N}$. Each interval can be expressed as $[t_{k-1}, \ t_k]$ for
$k = 1,...,N$. Denote the approximation of $u_h$ on the interval $[t_{k-1}, \
			t_k]$ by $u_k^h$, i.e. $u_h(t_k) = u_k^h$ for each $k$.\\

\sout{The goal the fully-discrete approximation as given by \cite{BV13} is to find an estimate between Problem GVar and Problem GVar$^h$, which can be expressed as}

@@

From Theorem 2 of the semi-discrete problem an error estimate for
$[u(t_k)-u_h(t_k)]$ with respect to the norm of $W$ was obtained. The authors
of \cite{BV13} then proceed to obtain an error estimate for $[u_h(t_k) -
			u^h_k]$. Since the dimension of $S^h$ is not fixed, the equivalence of norms
cannot be used, and therefore this error estimate for $[u_h(t_k) - u^h_k]$
should also be with respect to the norm of $W$.

\sout{In Theorem 2, Section 3.1.3 it is shown that \cite{BV13} obtained an error estimate for $[u(t_k)-u_h(t_k)]$ with respect to the norm of $W$. The authors explain that an estimate for $[u_h(t_k) - u^h_k]$ should also be with respect to the norm $W$ since the dimension of $S^h$ is not fixed and hence the equivalence of the relevant norms cannot be used.}\\

\subsubsection{Notation}
For any sequence $\{x_k\} \in R_n$,
For any sequence $\{x_k\} \in R_n$,
\begin{eqnarray*}
	\delta_{t}x_{k}& = &\tau^{-1}[x_{k+1}-x_{k}],\\
	x_{k+\frac{1}{2}} & = & \frac{1}{2}[x_{k+1}+x_{k}].
\end{eqnarray*}

Let Problem GVar$^h$-D denote the fully discrete form of Problem GVar$^h$.
\subsubsection*{Problem GVar$^\mathbf{{h}}$-D} Find a sequence
$\{u_{k}^{h}\}\subset S^{h}$ such that for $k=0,1,2$, . . . , $N-1,$
\begin{eqnarray}
	\delta_{t}u_{k}^{h}\ & = & v_{k+\frac{1}{2}},\\
	c(\delta_{t}v_{k},\ \psi)+a(v_{k+\frac{1}{2}},\ \psi)+b(u_{k+\frac{1}{2}}^{h},\ \psi) & = & \frac{1}{2}([f(t_{k})+f(t_{k+1})],\ \psi)_{X} \nonumber \\ \label{PP}
\end{eqnarray}
for each $\psi\in S^{h}$, while $u_{0}^{h}=u_{h}(0)=d^{h}$ and $v_{0}=u_{h}'(0)=v^{h}$.\\

\textcolor{red}{Haal uit************************************************}\\
The article \cite{BV13} proves the following proposition. This proposition proves that Problem GVar$^h$-D is well defined.

\newtheorem{DC_Prop4}[DC_Prop2]{Proposition}
\begin{DC_Prop4}
	Problem $GVar^{h}-D$ has a unique solution for any pair of vectors $d^{h}$ and $v^{h}$ in $S^{h}$.
\end{DC_Prop4}

In the proof of Proposition 3, the authors of \cite{BV13} show that \eqref{PP}
in Problem GVar$^h$-D can be rewritten as
\begin{eqnarray}
	c(v_k+1,\phi) + \frac{\tau}{2} a(v_{k+1},\phi) + \frac{\tau^2}{4} a(v_{k+1},\phi) & = & c(v_k,\phi) - \frac{\tau}{2} a(v_k,\phi) \nonumber \\  &&
	- \frac{\tau^2}{4}b(v_k,\phi) - \tau b(u_k,\phi) \nonumber\\
	&&+\frac{\tau}{2}([f(t_k)+f(t_{k+1})],\phi)_X.\nonumber \\ \label{PPP}
\end{eqnarray} for each $\phi \in S^h$.\\

Using \eqref{PPP} and the definition of the standard finite element matrices,
Problem GVar$^h$-D can be reformulated into a algorithm for the Finite Element
Method. This is the fully discrete form of Problem GVar and is represented in
Problem FD.\\
\textcolor{red}{******************************************************}\\

Further results and the standard Finite Element Matrices, let the authors of
\cite{BV13} rewrite Problem GVar$^\mathbf{{h}}$-D into an algorithm for the
Finite Element Method. This is represented as Problem FD

\subsubsection*{Problem FD}
Find a sequence $\{\bar{u}_{k}\}\subset R_{n}$ such that for each $k,$
\begin{eqnarray*}
	\bar{u}_{k+1}\ & = & \bar{u}_{k}+\tau\bar{v}_{k+\frac{1}{2}},\\
	(M + \frac{\tau}{2}C+\frac{\tau^{2}}{4}K)\bar{v}_{k+1} & = & (M- \frac{\tau}{2}C-\frac{\tau^{2}}{4}K)\bar{v}_{k}-\tau K\bar{u}_{k}+\frac{\tau}{2}[F(t_{k})+F(t_{k+1})]
\end{eqnarray*}
with $\bar{u}_{0}=\bar{d}$ and $\bar{v}_{0}=\bar{v}.$\\

In this form, the problem is fully discrete. This form also allow the authors
of \cite{BV13} to obtain an error estimate for the fully-discrete problem.

@@

\textcolor{red}{Haal uit*****************************}\\
There error estimate for $[u_h(t_k) - u^h_k]$ can now be determined. Recall that $\tau = \frac{T}{N}$. Subsituting $t = t_k$ and $t = t_k+1$ in Problem GVar$^h$ results in
\begin{eqnarray*}
	c( \tau^{-1}[v_{h}(t_{k+1})-v_{h}(t_{k})], \psi)+\frac{1}{2}a([v_{h}(t_{k+1})+v_{h}(t_{k})], \psi)+\frac{1}{2}b([u_{h}(t_{k+1})+u_{h}(t_{k})], \psi) \\
	=\frac{1}{2}([f(t_{k+1})+f(t_{k})], \psi)_{X}+c(\rho_{k}, \psi)
\end{eqnarray*}
where $v_{h}(t)=u_{h}'(t)$ and
\begin{eqnarray*}
	\rho_{k}=\tau^{-1}[v_{h}(t_{k+1})-v_{h}(t_{k})]-\frac{1}{2}[v_{h}'(t_{k+1})+v_{h}'(t_{k})].
\end{eqnarray*}

>>>>

To simplify the proof, the problem is divided into a semi-discrete problem and a fully discrete problem. The two terms are investigated as seperate problems
in \cite{BV13}.

@@

In \eqref{error_e}, the the error for the semi-discrete problem is the term $[u(t_k)-u_h(t_k)]$ and the term $[u_h(t_k) - u^h_k]$ is the error for the fully disrete appreximation of the semi-discrete approximation.

@@

\newtheorem*{DC_Thm5}{Theorem 2}
\begin{DC_Thm5}
	{Main Result}\\
	If $f\in C^{2}([0,T]4,X)$, then
	\begin{eqnarray*}
		\Vert u(t_{k})-u_{k}^{h}\Vert_{W}\ & \leq & 7T^{2}\tau^{2}\max\Vert u_{h}^{(4)}\Vert_{W}+7T\tau^{2}\max\Vert u_{h}'''\Vert_{W}\\
		& & + \sqrt{2K_{a}}\tau^{4}\max\Vert u_{h}'''\Vert_{W}
	\end{eqnarray*}
	for each $t\in(0,\ T)$.
\end{DC_Thm5}

@@

\subsection{Fully discrete Galerkin finite element approximation}
\textcolor{red}{********************HOU VOORLOPIG**********************}
Consider the time interval $J = [0,T]$. Divide $J$ into $N$ steps with length
$\tau = \frac{T}{N}$. Each interval can be expressed as $[t_{k-1}, \ t_k]$ for
$k = 1,...,N$. Denote the approximation of $u_h$ on the interval $[t_{k-1}, \
			t_k]$ by $u_k^h$, i.e. $u_h(t_k) = u_k^h$ for each $k$.\\

@@

From Theorem 2 of the semi-discrete problem an error estimate for
$[u(t_k)-u_h(t_k)]$ with respect to the norm of $W$ was obtained. The authors
of \cite{BV13} then proceed to obtain an error estimate for $[u_h(t_k) -
			u^h_k]$. Since the dimension of $S^h$ is not fixed, the equivalence of norms
cannot be used, and therefore this error estimate for $[u_h(t_k) - u^h_k]$
should also be with respect to the norm of $W$.
\textcolor{red}{********************HOU VOORLOPIG**********************}

@@

\textcolor{red}{Moontlik hou!}\\
There error estimate for $[u_h(t_k) - u^h_k]$ can now be determined. Recall that $\tau = \frac{T}{N}$. Subsituting $t = t_k$ and $t = t_k+1$ in Problem GVar$^h$ results in
\begin{eqnarray*}
	c( \tau^{-1}[v_{h}(t_{k+1})-v_{h}(t_{k})], \psi)+\frac{1}{2}a([v_{h}(t_{k+1})+v_{h}(t_{k})], \psi)+\frac{1}{2}b([u_{h}(t_{k+1})+u_{h}(t_{k})], \psi) \\
	=\frac{1}{2}([f(t_{k+1})+f(t_{k})], \psi)_{X}+c(\rho_{k}, \psi)
\end{eqnarray*}
where $v_{h}(t)=u_{h}'(t)$ and
\begin{eqnarray*}
	\rho_{k}=\tau^{-1}[v_{h}(t_{k+1})-v_{h}(t_{k})]-\frac{1}{2}[v_{h}'(t_{k+1})+v_{h}'(t_{k})].
\end{eqnarray*}

### 2023-11-11 files/Overview.tex
Overview, Chapter 3

\subsubsection{Chapter 1}
This dissertation is a literature study to investigate the validity of different linear models for beams and plates. The term validity in this dissertation means how well a model compares to a more realistic model for real world applications. The first chapter of the dissertation introduces the models. The simplest model is the Timoshenko model. The other models are a two-dimensional elastic beam model, three-dimensional elastic beam and plate models and a Reissner-Mindlin plate model. The aim in this dissertation to validate the use of the Timoshenko beam model and the Reissner-Mindlin plate model in applications. These models are simplified one-dimensional beam and two-dimensional plate models. But more realistic models exists, such as a two-dimensional and three-dimensional beam model and a three-dimensional plate model. Using modal analysis (see Chapter 2), it is shown that the solutions of the models can be represented as a linear combination of the modal solutions. Using this idea, it is only required to compare the eigenvalues and eigenfunctions of the models to be able to compare the difference between the solutions. In Chapter 1, the models of the dissertation is given (they are given in a dimensionless form).  Model problems are then defined by including boundary conditions. These model problems are the models used in the rest of the dissertation. The variational forms are also derived which enables us to see the similarity between the models and the general theory.

\subsubsection{Chapter 2}
First, a variational form for a model problem of a cantilever Timoshenko beam given. This model problem in variational form is then extended to complete function spaces, and a weak variational problem is obtained. This weak variational problem is used as an example to explain the main theory of this chapter which is from the article \cite{VV02}. In this article, the authors present a general weak variational form. The weak variational form of all the model problems of this dissertation are special cases of this general weak variational problem. Therefore the assumptions and results can be formulated and applied to all the applications in this dissertation. The article gives four assumptions. Under these assumptions it is shown that the general vibration problem has a unique solution using semi-group theory. The theory is then applied to the example by proving that the assumptions hold.

Finally the concept of modal analysis is introduced, which is fundamental to this dissertation. First idea of modal analysis is explained by hands of another example, again using a cantilever Timoshenko beam. Then the general case is discussed that follows the article \cite{CVV18}. Given a general vibration problem, it is split into two problems by introducing a trial solution. It can then be verified that these two problems are the eigenvalue problem and a differential equation. An additional assumption (additional to the assumptions of \cite{VV02}) is introduced by \cite{CVV18}. Under this assumption, the eigenvalue problem has infinite many solutions, and the solution of the general vibration problem is an infinite series of the eigenvectors and general Fourier coefficients. Solving the differential equation, shows that these Fourier coefficients are periodic and a formal series solution for the general vibration problem (which is a boundary value problem) is obtained. This formal series solution is then shown to also be valid for a initial value problem. This general case is also applicable to all the models in this dissertation.

@@

\subsubsection{Chapter 3}
\textcolor{red}{Nie seker hoekom prof gese het nie BV13 nie? Ook nie seker hoekom SF97 dood getrek was nie.}

In this chapter, some theory for the Finite Element Method is discussed. This chapter contains two parts, that covers two different theoretical results. The first is on the convergence of the Galerkin approximation of a general vibration problem. This theory is presented in the article \cite{BV13}. The second part concerns the convergence of the eigenvalues and eigenfunctions of a one-dimensional vibration problem when applying the Finite Element Method. This theory is presented in the textbook \cite{SF97}.

In the first section, the authors continue with the general vibration problem that is used in Chapter 2. The general vibration problem has a solution as shown in Chapter 2. The Galerkin Approximation is derived from the general vibration probelm and is rewritten into a system of ordinary differential using the Finite Element Method. This oridinary differential equation can be proven to have a unique solution. The main results of \cite{BV13} shows that this solution of the Galerkin Approximation converges to the solution of the general vibration problem. The approach of the authors is to calculate the error estimates. The error is split into two parts, a semi-discrete and a fully discrete problem. Two assumptions are added to the assumptions from Chapter 2. Using these assumptions, the authors obtain error estimates for the semi-discrete and fully discrete problems.

The second part of the chapter considers work done in a textbook \cite{SF97}. The specific work discussed, covers the convergence of the eigenvalues and eigenfunctions of a one-dimensional vibration problem when applying the Finite Element Method. The authors consider a general eigenvalue problem. Then using the Rayleigh-quotient, from the Rayleigh-Ritz method, as well as an approximation theorem from \cite{OR97}, the main result is proven. The specific work done in this section is updating the notation of the textbook, as well as expanding some results so that the results are easier to understand.

\subsubsection{Chapter 4}
This chapter is a focus on the main model theory of this dissertation, the Timoshenko beam theory. The first section is a discussion of modal analysis applied to the Timoshenko beam theory, and specifically a discussion of the article \cite{VV06}. In this article, the authors present a method to calculate the exact eigenvalues and eigenfunctions of a Timoshenko beam. This method considers the boundary conditions, which differs from most other research. Starting with a general eigenvalue problem for a Timoshenko beam model, the authors derive a general solution for this problem. The authors then explain the method by hands of an example by applying the method to a cantilever beam model. The next sections of this chapter also then looks at examples of applying the method, first to a cantilever beam model, and then to a free-free beam model. The examples shows the use of the method, and is preperation for chapter 6 where these results are required for the comparison of the models.

The next section is a short discussion on a suspended beam model, afterwards an important article \cite{SP06} is discussed. In this article, the authors investigate the validity of the Timoshenko beam theory, by comparing the eigenvalues (natural frequencies) for a physical beam, to a Timoshenko beam and a three-dimensional beam using Finite Element Analysis. The authors suspended the beam, and vribrated the beam to induce a free-free beam and measure the natural frequencies. These emperical results are then compared to the theoretical results. This result, together with the results in chapter 6, give a good picture to the validity of the Timoshenko beam theory. Finally the chapter finished by looking at an interesting model of a cantilever beam with a tip-body with elastic interfaces. These results combine modal analysis and investigates at the behavior of the model. This section is an example of the usefullness of the Timoshenko beam theory. A similar model using realistic two- or three-dimensional model will be very challenging.

>>>>

\subsubsection{Chapter 1}
This dissertation is a literature study to investigate the validity of different linear models for beams and plates. The term validity in this dissertation means how well a model compares to a more realistic model for real world applications. The first chapter of the dissertation introduces the models. The simplest model is the Timoshenko model. The other models are a two-dimensional elastic beam model, three-dimensional elastic beam and plate models and a Reissner-Mindlin plate model. The aim in this dissertation to validate the use of the Timoshenko beam model and the Reissner-Mindlin plate model in applications. These models are simplified one-dimensional beam and two-dimensional plate models. But more realistic models exists, such as a two-dimensional and three-dimensional beam model and a three-dimensional plate model. Using modal analysis (see Chapter 2), it is shown that the solutions of the models can be represented as a linear combination of the modal solutions. Using this idea, it is only required to compare the eigenvalues and eigenfunctions of the models to be able to compare the difference between the solutions. In Chapter 1, the models of the dissertation are given (in dimensionless form).  Model problems are then defined by including boundary conditions. These model problems are the models used in the rest of the dissertation. The variational forms are also derived which enables us to see the similarity between the models and the general theory.

\subsubsection{Chapter 2}
First, a variational form for a model problem of a cantilever Timoshenko beam is given. This model problem in variational form is then extended to complete function spaces, and a weak variational problem is obtained. This weak variational problem is used as an example to explain the main theory of this chapter which is from the article \cite{VV02}. In this article, the authors present a general weak variational form. The weak variational form of all the model problems of this dissertation are special cases of this general weak variational problem. Therefore the assumptions and results can be formulated and applied to all the applications in this dissertation. The article gives four assumptions. Under these assumptions it is shown that the general vibration problem has a unique solution using semi-group theory. The theory is then applied to the example by proving that the assumptions hold.

Finally the concept of modal analysis is introduced, which is fundamental to this dissertation. First idea of modal analysis is explained by hands of another example, again using a cantilever Timoshenko beam. Then the general case is discussed that follows the article \cite{CVV18}. Given a general vibration problem, it is split into two problems by introducing a trial solution. It can then be verified that these two problems are the eigenvalue problem and an ordinary differential equation. An additional assumption (additional to the assumptions of \cite{VV02}) is introduced by \cite{CVV18}. Under this assumption, the eigenvalue problem has a complete orthonormal sequence of eigensolutions. The solution of the general vibration problem is an infinite series of modal solutions. This formal series solution is then shown to also be valid for a initial value problem. This general case is also applicable to all the models in this dissertation.

@@

\subsubsection{Chapter 3}
In this chapter, some theory for the Finite Element Method is discussed. This chapter contains two parts, that covers two different theoretical results. The first is on the convergence of the Galerkin approximation of a general vibration problem. This theory is presented in the article \cite{BV13}. The second part concerns the convergence of the eigenvalues and eigenfunctions of a vibration problem when applying the Finite Element Method. This theory is presented in the textbook \cite{SF97}\textcolor{red}{Nie goed genoeg vir dinamiese probleem}.

In the first part, the authors of \cite{BV13} consider the general vibration problem that is studied in Chapter 2. The general vibration problem has a solution as shown in Chapter 2. The Galerkin Approximation is derived from the general vibration problem and is rewritten into a system of ordinary differential using the Finite Element Method. This oridinary differential equation can be proven to have a unique solution. The main results of \cite{BV13} shows that this solution of the Galerkin Approximation converges to the solution of the general vibration problem. The approach of the authors is to calculate the error estimates.

The second part of the chapter considers work done in a textbook \cite{SF97} on eigenvalue problems for elliptic partial differential equations. The specific work discussed, covers the convergence of the eigenvalues and eigenfunctions of a one-dimensional vibration problem when applying the Finite Element Method. The authors consider a general eigenvalue problem. Then using the Rayleigh-quotient, from the Rayleigh-Ritz method, as well as an approximation theorem from \cite{OR97}, the main result is proven. The specific work done in this section is updating the notation of the textbook, as well as expanding some proofs so that the results are easier to understand.

\subsubsection{Chapter 4}
This chapter is a focus on the main theory of this dissertation, the Timoshenko beam theory. The first section is a discussion of modal analysis applied to the Timoshenko beam theory, and specifically a discussion of the article \cite{VV06}. In this article, the authors present a method to calculate the exact eigenvalues and eigenfunctions of a Timoshenko beam. Starting with a general eigenvalue problem for a Timoshenko beam model, the authors derive a general solution for the ordinary differential equation. The authors then explain the method by hands of an example by applying the method to a cantilever beam model. The next sections of this chapter also then looks at examples of applying the method, first to a cantilever beam model, and then to a pinned-pinned beam model.

The next section is a short discussion on a suspended beam model, from an important article \cite{SP06}. In this article, the authors investigate the validity of the Timoshenko beam theory, by comparing the eigenvalues (natural frequencies) for a physical beam, to a Timoshenko beam and a three-dimensional beam using Finite Element Analysis. The authors report on an experiment with a vibrated a free-free beam measure the natural frequencies. These emperical results are then compared to the theoretical results. This result, together with the results in chapter 6, give a good picture to the validity of the Timoshenko beam theory. Finally the chapter is concluded by looking at a model of a cantilever beam with a tip-body with elastic interfaces. These results combine modal analysis and investigates at the behavior of the model. This section is an example of the usefullness of the Timoshenko beam theory. A similar model using realistic two- or three-dimensional model will be very challenging.
